---
title: "Exploring How Napping Affects Infant Cognitive Development"
author: "Albert Sebastian"
date: "12-04-2024"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(ggplot2)
library(dplyr)
data <- read_csv("infants_data_DI_task.csv")

```

## Research Questions

-   Big Research Question: How do environmental conditions and caregiver involvement impact infants' behavioral performance and physiological processes such as sleep quality and imitation skills?

-   Little research Question: What is the effect of napping on infants' cognitive task performance without parental assistance?

## Method

### Participants

The participants in this study are infants from two different age groups: 15 months and 24 months. There were a total of 105 participants.

### Study Design

Participants were asked to expose themselves to media (tablet or picture book) prior to bedtime 3 days in a row. They were then asked to engage in a series of tasks where infants either imitated actions with or without caregiver support, as part of a deferred imitation task.

## Plot

```{r Data Prep, echo=FALSE}

processed_data <- data %>%
  mutate(NoScaff_LTFA_Score = ifelse(is.na(NoScaff_LTFA_Score), 0, NoScaff_LTFA_Score))%>%
  group_by(condition) %>%
  summarise(
    target_actions = mean(NoScaff_Imi_Score[NoScaff_Imi_Score != 0], na.rm = TRUE),
    latency_to_first_action = mean(NoScaff_LTFA_Score[NoScaff_LTFA_Score != 0], na.rm = TRUE)
  ) %>%
  pivot_longer(cols = -condition, names_to = "metric", values_to = "average_score")

```

```{r plot, echo=FALSE}
ggplot(processed_data, aes(x = metric, y = average_score, fill = condition)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(
    title = "Effect of Napping on Imitation Task Performance (No Scaffolding)",
    x = "",
    y = "Average Score",
    fill = "Condition"
  ) +
  scale_x_discrete(labels = c("Latency to First Action", "Target Actions")) +
  theme_minimal()
```
