---
title: "D2M Assignment Notebook"
output: html_notebook
author: "Albert Sebastian"
---
```{r}
library(tidyverse)
```


#ICMA 9/11/2024

Variable names cannot have special symbols
Error: 0hno!This_is.a_nightmare <- "hello"
Fixed: message<- "hello"

2+2 in console: Output shows up in console window
2+2 in a  chunk: Output shows up right below the  chunk

x <- 2 + 2 in a code chunk: No output
y <- 2 + 2 in console: No output
In environment pane, x and y are loisted as variables with values of 4

?mean: R displays a help page for the mean function
help(mean): Same output as ?mean

```{r}
data <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10)
result <- sd(data)
print(result)
```
#ICMA 9/16/2024

```{r}
my_numbers <- c(1, 2, 3, 4, 5, 8, 9, 10)
mean_value <- mean(my_numbers)
print(mean_value)
mean_value_pipe <- my_numbers %>% mean()
print(mean_value_pipe)
```
The two outputs are the same

```{r}
?mtcars
write_csv(mtcars, "mtcars_fromR.csv")
read_csv("mtcars_fromR.csv")
```
Everything appears as expected with no missing values or empty columns/rows

#ICMA 9/18/2024

```{r}
chick_clean <- ChickWeight %>% ## assigning to a new object "chick_clean"
  select(chick_id = Chick, 
         diet = Diet,
         age_days = Time, 
         weight_gm = weight)%>%  ## renaming & re-ordering while keeping
  mutate(weight_lb = weight_gm / 453.6)  # Creating a new column for weight in pounds

# Create a summary table for Day 20
summary_table <- chick_clean %>%
  filter(age_days == 20) %>%  # Filter for Day 20
  group_by(diet) %>%  # Group by diet
  summarise(
    num_chicks = n(),  # Count of chicks
    avg_weight_gm = mean(weight_gm, na.rm = TRUE),  # Average weight in grams
    avg_weight_lb = mean(weight_lb, na.rm = TRUE)  # Average weight in pounds
  )

# Print the summary table
print(summary_table)
```




