---
title: "D2SC: Weekly Assignments"
output: html_notebook
author: "Albert Sebastian"
date: 'r Sys.Date()'
---

# Initial Loading

```{r}
library(tidyverse)
```


# Weekly Assignment 1

To figure out how to load a package in R, I started by searching for "how to load an R package" in Google. I found several helpful resources, including Stack Overflow and the RStudio website. One useful link was 'Stack Overflow - How to Load a Package in R' (https://stackoverflow.com/questions/11805142/how-to-load-a-package-in-r). 

I verified that the 'tidyverse' package was correctly loaded by running 'library(tidyverse)' in the R console and checking that no error messages appeared. I also used the 'sessionInfo()' function to confirm that 'tidyverse' was listed among the loaded packages.

```{r}
help(package = "tidyverse")
```

The 'tidyverse' is a set of packages that work in harmony because they share common data representations and 'API' design. This package is designed to make it easy to install and load multiple 'tidyverse' packages in a single step. Learn more about the 'tidyverse' at https://www.tidyverse.org.


# Weekly Assignment 2

```{r}
library(readr)
analogy_data <- read_csv("\\Users\\alber\\Documents\\GitHub\\CogSci412-Data2SciComm\\WeeklyAssignments\\WeeklyAssignments\\tidy_data\\MFIndD_analogy.csv")
analogy_data

```
The qualtrics_id column seems to be the unique identifier for each participant

```{r}
library(dplyr)

dimensions <- dim(analogy_data)
rows <- dimensions[1]
cols <- dimensions[2]
rows
cols
```
2a. There are 792 rows and 6 columns in the dataset. The dim() function was used to retrieve the dimensions of the dataset, specifically the number of rows and columns. By calling dim(analogy_data), it returned a vector where the first element indicated the number of rows and the second element indicated the number of columns.
```{r}
unique_participants <- analogy_data %>%
  distinct(qualtrics_id) %>%
  count()
unique_participants
```
There are 99 unique people in the dataset.
```{r}
trials_per_participant <- analogy_data %>%
  group_by(qualtrics_id) %>%
  summarise(trial_count = n())

consistent_trials <- n_distinct(trials_per_participant$trial_count) == 1
consistent_trials
```
Everyone has the same number of trials in this dataset

# Weekly Assignment 3
```{r, eval=FALSE}
# Attempting to summarize relational matches
summary_data <- analogy_data %>%
  group_by(participant_id) %>%
  summarize(relational_matches = sum(response_type == "Relational", na.rm = TRUE))
```

In this chunk, I mistakenly used "Relational" as the condition in the 'sum()' function instead of "Rel." As a result, the summarization returned 'NA' values because there were no matches found for that string in the 'response_type' column. 

To fix it, I:
1. Reviewed the dataset to check the exact labels used for the responses.
2. Realized that the correct label was `"Rel"` for relational matches.
3. Updated the code accordingly:

```{r}
# Summarize the data
summary_data <- analogy_data %>%
  group_by(qualtrics_id) %>%
  summarize(relational_matches = sum(response_type == "Rel", na.rm = TRUE))

# Check the result
head(summary_data)
```

```{r}
library(ggplot2)

# Create a histogram
ggplot(summary_data, aes(x = relational_matches)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Relational Matches",
       x = "Number of Relational Matches",
       y = "Frequency") +
  theme_minimal()
```
The histogram shows a right-skewed distribution, with most participants selecting a low number of relational matches, and a few participants achieving higher counts.This indicates that the majority of participants may struggle with recognizing relational matches in the trials.

```{r}
# Reshape the data
reshaped_data <- analogy_data %>%
  select(qualtrics_id, trial_number, response_type) %>%
  pivot_wider(names_from = trial_number, values_from = response_type, names_prefix = "Trial_")

# Check the reshaped data
head(reshaped_data)
```

# Weekly Assignment 4


```{r}
library(dplyr)

data <- read.csv("WeeklyAssignments\\tidy_data\\MFIndD_REI.csv")  

# Preview the data
head(data)

# Check column types
str(data$response)
str(data$scored_response)
```
The "response" column is of type character while the "scored_respose" column is a numeric(int) type.The response column contains the actual survey responses which is not inherently numeric, so it is stored as a character or factor type. The scored_response column contains numeric values that have been derived from the responses, specifically designed to quantify the survey responses for analysis.

```{r}

# Convert the response column into numeric values
data <- data %>%
  mutate(response_numeric = as.double(case_when(
    response == "Strongly Disagree" ~ 1,
    response == "Disagree" ~ 2,
    response == "Neutral" ~ 3,
    response == "Agree" ~ 4,
    response == "Strongly Agree" ~ 5,
    TRUE ~ NA_real_  # Handling any unexpected values
  )))

data$response_numeric[is.na(data$response_numeric)] <- as.double(data$response[is.na(data$response_numeric)])

head(data)
```


```{r}
data <- data %>%
  mutate(new_scored_response = case_when(
    is.na(rev_scoring) ~ response_numeric,
    TRUE ~ 6 - response_numeric  # Reverse scoring logic
  ))

head(data)
```



```{r}
# Compare new_scored_response with scored_response
comparison <- data %>%
  select(new_scored_response, scored_response) %>%
  mutate(match = new_scored_response == scored_response)

# Summary of matches
match_summary <- comparison %>%
  summarise(
    total_rows = n(),
    matches = sum(match, na.rm = TRUE),
    mismatch_count = total_rows - matches
  )

print(match_summary)
```
# Weekly Assignment 5
```{r}
library(dplyr)
library(readr)

# Read the dataset
data <- read_csv("WeeklyAssignments/tidy_data/MFIndD_REI.csv")

# Create a new column for reverse scoring
data <- data %>%
  mutate(new_scored_response = case_when(
    is.na(rev_scoring) ~ scored_response,
    TRUE ~ 6 - scored_response  # Reverse scoring logic
  ))

# Summarize the data to get one row per participant and sub_scale
summary_data2 <- data %>%
  group_by(qualtrics_id, sub_type) %>%  # Group by participant and sub_scale
  summarise(score = sum(new_scored_response, na.rm = TRUE), .groups = 'drop')

# Preview the summarized data
print(head(summary_data2))



```



```{r}
# Check for NA scores in the summarized data
na_scores <- summary_data2 %>%
  filter(is.na(score))

# Count the number of NA scores
na_count <- nrow(na_scores)

# Output the count of NA scores
na_count

```
The code filters the summary_data for any rows where the score is NA and counts how many such rows exist. Since na_count returns 0, there are no NA scores in the summary.


```{r}
# Summarize the analogy dataset
analogy_summary <- analogy_data %>%
  group_by(qualtrics_id) %>%
  summarize(relational_matches = sum(response_type == "Rel", na.rm = TRUE), .groups = 'drop')

# Combine the two summarized datasets
combined_data <- full_join(summary_data2, analogy_summary, by = "qualtrics_id")

# Preview the combined dataset
print(head(combined_data))



```

```{r}
library(ggplot2)

# Create the scatterplot
ggplot(combined_data, aes(x = score, y = relational_matches, color = sub_type)) +
  geom_point(size = 3, alpha = 0.7) +  # Scatter points
  geom_smooth(method = "lm", se = TRUE, linetype = "dashed", color = "black") +  # Smoothed line
  labs(title = "Relationship Between REI Score and Analogy Score",
       x = "REI Score",
       y = "Analogy Score (Relational Matches)",
       color = "REI Sub-Type") +  # Custom axis labels
  theme_minimal() +  # Clean theme
  theme(panel.grid.major = element_line(color = "lightgray"),  # Change grid line color
        panel.grid.minor = element_blank())  # Remove minor grid lines

```

Based on the plot, I noticed that there is not a strong correlation between the two values as modeled by the best fit function being close to horizontal. Each REI sub-type has more data points with an analogy score of 4 or less except for type EE which has more data points with an analogy score of 6 or more.











