---
title: "D2SC: Weekly Assignments"
output: html_notebook
author: "Albert Sebastian"
date: "r Sys.Date()"
---

# Initial Loading

```{r}
library(tidyverse)
```


# Weekly Assignment 1

To figure out how to load a package in R, I started by searching for "how to load an R package" in Google. I found several helpful resources, including Stack Overflow and the RStudio website. One useful link was 'Stack Overflow - How to Load a Package in R' (https://stackoverflow.com/questions/11805142/how-to-load-a-package-in-r). 

I verified that the 'tidyverse' package was correctly loaded by running 'library(tidyverse)' in the R console and checking that no error messages appeared. I also used the 'sessionInfo()' function to confirm that 'tidyverse' was listed among the loaded packages.

```{r}
help(package = "tidyverse")
```

The 'tidyverse' is a set of packages that work in harmony because they share common data representations and 'API' design. This package is designed to make it easy to install and load multiple 'tidyverse' packages in a single step. Learn more about the 'tidyverse' at https://www.tidyverse.org.


# Weekly Assignment 2

```{r}
library(readr)
analogy_data <- read_csv("\\Users\\alber\\Documents\\GitHub\\CogSci412-Data2SciComm\\WeeklyAssignments\\WeeklyAssignments\\tidy_data\\MFIndD_analogy.csv")
analogy_data

```
The qualtrics_id column seems to be the unique identifier for each participant

```{r}
library(dplyr)

dimensions <- dim(analogy_data)
rows <- dimensions[1]
cols <- dimensions[2]
rows
cols
```
2a. There are 792 rows and 6 columns in the dataset. The dim() function was used to retrieve the dimensions of the dataset, specifically the number of rows and columns. By calling dim(analogy_data), it returned a vector where the first element indicated the number of rows and the second element indicated the number of columns.
```{r}
unique_participants <- analogy_data %>%
  distinct(qualtrics_id) %>%
  count()
unique_participants
```
There are 99 unique people in the dataset.
```{r}
trials_per_participant <- analogy_data %>%
  group_by(qualtrics_id) %>%
  summarise(trial_count = n())

consistent_trials <- n_distinct(trials_per_participant$trial_count) == 1
consistent_trials
```
Everyone has the same number of trials in this dataset

# Weekly Assignment 3
```{r, eval=FALSE}
# Attempting to summarize relational matches
summary_data <- analogy_data %>%
  group_by(participant_id) %>%
  summarize(relational_matches = sum(response_type == "Relational", na.rm = TRUE))
```

In this chunk, I mistakenly used "Relational" as the condition in the 'sum()' function instead of "Rel." As a result, the summarization returned 'NA' values because there were no matches found for that string in the 'response_type' column. 

To fix it, I:
1. Reviewed the dataset to check the exact labels used for the responses.
2. Realized that the correct label was `"Rel"` for relational matches.
3. Updated the code accordingly:

```{r}
# Summarize the data
summary_data <- analogy_data %>%
  group_by(qualtrics_id) %>%
  summarize(relational_matches = sum(response_type == "Rel", na.rm = TRUE))

# Check the result
head(summary_data)
```

```{r}
library(ggplot2)

# Create a histogram
ggplot(summary_data, aes(x = relational_matches)) +
  geom_histogram(binwidth = 1, fill = "blue", color = "black") +
  labs(title = "Distribution of Relational Matches",
       x = "Number of Relational Matches",
       y = "Frequency") +
  theme_minimal()
```
The histogram shows a right-skewed distribution, with most participants selecting a low number of relational matches, and a few participants achieving higher counts.This indicates that the majority of participants may struggle with recognizing relational matches in the trials.

```{r}
# Reshape the data
reshaped_data <- analogy_data %>%
  select(qualtrics_id, trial_number, response_type) %>%
  pivot_wider(names_from = trial_number, values_from = response_type, names_prefix = "Trial_")

# Check the reshaped data
head(reshaped_data)
```

